---
title: <center>Historical Biogeography</center>
author: <center>Jeet Sukaumaran</center>
date: <center>`r format(Sys.Date(),format="%%B %%d %%Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

May move back and forth with Textmate
```{r}
library(optimx)
library(BioGeoBEARS)
```
Continuous vs discrete approaches - latter is appropriate when you can focus on elements of interest. Biogeo - species level and higher, where discrete applies.

DEC - Dispersal-Extinction-Cladogenesis model - range evolution - adding and removing areas (local extinction - a lineage leaves an area).  Leads to inference of area of origin for lineage.

#### Input

Phylogeny and Geography

* ultrameric tree
* geographical distribution of tips

Phylogeny is accepted as true.

#### Output

* Likelihood of model
* Estimation of global extinction dispersal rates
* Likelihood of ranges.

Ranges- consist of a set of occupied areas and constitute character states.  Evolve either anagenically or cladogenically.

Dispersal - product of global rate d and connectivity of two areas involved (i ->j) [relative - 0 to 1]

Uncomment if needed.
```{r}
# setwd("./ssb2015-workshop/2015-SSB-AnnArbor-Biogeography/biogeobears")
```
Note that establishing ranges for d and e are critical 

dec.j adds jump parameter to estimate - mimics island colonization.
