---
title: "SSB Workshop on Comparative Methods"
author: "Bruce Cochrane"
date: "May 5, 2015"
output: html_document
---


Code from Brian O'Meara
```{r}
#install.packages("ctv")
```
```{r}
library(ctv)
#install.views("Phylogenetics")
```
```{r}
x <-available.views()
print(x,packagelist=TRUE)
```
list the names of the views
```{r}
sapply(x, function(y) y$name)
```
Ok.  so the phylogeny package is element 23 on the list
```{r}
phylo <-x[[23]]
str(phylo)
```
To get the list

```{r}
phylo.pkg <-phylo$packagelist
phylo.pkg
```
And Genetics is number 11.  We can list the packages
```{r}
gpks <-x[[11]]$packagelist
gpks
```


We should write this to a file so we can print it

```{r}
#write.csv(phylo.pkg,file="./Phyloview.pkg.csv",quote=FALSE,row.names=FALSE)
```
Here's a fun one - Paleotree
```{r}
library(paleotree)
```
```{r}
taxa <- simFossilTaxa(p=0.1,q=0.05,nruns=1,mintaxa=20,maxtaxa=30,maxtime=1000,maxExtant=50,plot=TRUE)
taxa
```
OK.  that generates a curve of taxon richness over time.  We need to figure out the units, but it's pretty cool (p=origination, q=extinction)

OK.  And now we get to imperfect sampling:

```{r}
rangesCont <-sampleRanges(taxa,r=.1)
taxicDivCont(rangesCont)
```
Now if the data are binned:

```{r}
rangesDisc <- binTimeData(rangesCont,int.length=10)
#plot with taxicDivDisc
taxicDivDisc(rangesDisc)
```
And that's the real key.

Back to ctv, The odd thing is that they don't show on the RStudio package list.  However, using installed.packages() they all show up.
