---
title: "DiagrammeR"
author: "Bruce Cochrane"
date: "May 1, 2015"
output: html_document
---
```{r}
#install.packages("DiagrammeR")
library(DiagrammeR)
```
### The 206 Concept map

Code below generates a slight revision of the original concept map.
```{r}
DiagrammeR("graph TB;
          b(Broad Evolutionary Question)-->Question;
           Question(Specific Question of Paper)-->BG1(Background 1);
           Question-->BG2(Background 2);
           Question-->BG3(Background 3);
           BG1-->Design/Methods(Experimental Design and Methods);
           BG2-->Design/Methods;
           BG3-->Design/Methods;
           Design/Methods-->Results1;
           Design/Methods-->Results2;
           Design/Methods-->Results3;
           Results1-->Conclusions;
           Results2-->Conclusions;
           Results3-->Conclusions;
           Conclusions-->OpenQuestions(Open Questions);
           OpenQuestions-->A(Full Citation);
          classDef default fill:#FC1,stroke:#333,stroke-width:2px;")
```
OK.  So the basics are clear.  We need to use the parenthesis notation to get it the best we want.  And another little trick - shape of box can be determined - () rounded, [] square, and {} diamond.  What would be nice would be if the default color could be set.  That is now done; information is available via [mermaid.js help](http://knsv.github.io/mermaid/) pages.  don't need to worry about the graphviz stuff for now.

### Make a pedigree

```{r}
library(DiagrammeR)
DiagrammeR('graph TB;
           A[A]-->C((C));
            B((B))-->C;')
  

```

