---
title: <center>From the Nescent Hackathon</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

```{r}
library(devtools)
install_github("georgeshirreff/multiNe")
```
```{r}
install.packages("INLA", repos="http://www.math.ntnu.no/inla/R/stable")
```

So that gets it done.

```{r}
library(multiNe)
```
```{r}
data(nancycats)
HENe(nancycats)
```
OK, so that computes Ne based on heterozygote excess.  It uses genind object (adegenet)

```{r}
summary(nancycats)
```
I think there's a vignette in pegas that may help with the genind construction - 

Another cool thing - generate simulated microsatellite data from ms run:

```{r}
ms.out <- ms(nsam=30,nreps=5,opts=("-s 10"))
ms.ms <-ms2ms(ms.out)
```
```{r}
summary(ms.ms)
```
Now, does it work?
```{r}
HENe(ms.ms)
```

A couple of intriguing possibilities

1.  Adegenet can read genepop files, so we can generate them in Genalex
2. Read hapmap data with chopsticks, getting snpmatrix; can then (we hope) convert to genind and run.

```{r}
ms2ms
```

