---
title: <center>Pedigree Drawing</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

 the package kinship2
```{r}
#install.packages("kinship2")
```
```{r}
library(kinship2)
```
First, use data supplied to get format
```{r}
data(minnbreast)
head(minnbreast)
```

So I wonder if we could create a dataframe via excel that just has the first four columns (actually five, including sex)

```{r}
dat.raw <-read.csv("./DATA/SimplePedigree.csv",stringsAsFactors=FALSE)
head(dat.raw)
```
Now create a pedigree object
```{r}
dat.ped <-with(dat.raw,pedigree(id,fatherid,motherid,sex))
plot(dat.ped)
```
That is super cool!!  We can use it in genetics next fall for all types of stuff.  And for just drawing a structure, all we need is three columns - motherid, fatherid and sex.  We're going to make a file for that
```{r}
dat.raw <-dat.raw[,c(1,3,4,6)]
dat.raw
```
```{r}
save(dat.raw,file="../Data2015/Pedigree")
```
### May 8

OK.  now we want to add in affected status.  We'll just do it randomly for now
```{r}
aff <-rbinom(12,1,.5)
aff
```
And we'll cbind that to the data
```{r}
dat.new <-cbind(dat.raw,aff)
head(dat.new)
```
And now we should be able to generate the file
```{r}
ped.2 <-with(dat.new, pedigree(id,fatherid,motherid,sex,aff))
plot(ped.2)
```
Well what do you know?  That actually works as an autosomal dominant!!  Of course, we didn't set.seed, so we'll never get it again . . . 


Try it with names
```{r}
dat.me <-read.csv("./Data/MyFamily.csv")
dat.me
```
```{r}
ped.me <-with(dat.me,pedigree(id,fatherid,motherid,sex,missid="0"))
plot(ped.me)
```

