---
title: <center>Phasing Haplotypes</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---
A start at phasing based on r hapmap data.

What we'll do is to try to get 10 snps from 1 population into a matrix and then use it.  I think the procedure will be 

1.  Read some data
2.  Genindify it.
3.  loci-ify it
4.  Modify the STRUCTURE script to generate a matrix
5.  Run the haplo.em function

```{r}
library(haplo.stats)
library(biomaRt)
```
```{r}
listMarts()
```
```{r}

hap <-useMart("HapMap_rel27")
```
```{r}
listDatasets(hap)
```
```{r}
mart.yri <-useMart("HapMap_rel27", dataset="hm27_variation_yri")
```

```{r}
yri.dat <-getBM(attributes=c("yri_id","marker1"),filters=c("chrom","start","stop"),values=list("chr3",10000000,10050000),mart=mart.yri)
```
That's good.  45 snps.  Now we need to to the rest
```{r}
colnames(yri.dat)
head(yri.dat)
```

So the first we'lll do is make genos as a convenience variable

```{r}
genos <-(yri.dat$yri_id)
```
Now gsub out the spaces
```{r}
genos <-lapply(genos, function (x) gsub(" ","",x))
genos[[1]]
```
OK.  Now to split them
```{r}
genos <-lapply(genos,strsplit,NULL)
genos[[1]]
```
Let's make the data frame now


```{r}
genos.tmp <-sapply(genos,unlist)
genos.df <-data.frame(t(genos.tmp))
```
Now, just out of curiosity, we can try the imputation
```{r}
genos.hap <-haplo.em(genos.df[,1:10],miss.val="N")
```

