---
title: "rvest"
author: "Bruce Cochrane"
date: "May 23, 2015"
output: html_document
---
From R Bloggers
```{r}
install.packages("rvest")
library(rvest)
```

Just use the sample code for now.
```{r}
URL <- "http://ultrasignup.com/entrants_event.aspx?did=31114"
rvest_doc <-html(URL)
```

```{r}
str(rvest_doc)
```
Then you need to find a node - does so via css.  in the example, this is done with "table.ultra_grid", but I suspect that is site-specific.  I wonder what would happen if we just used "table"
```{r}
test2 <-html_nodes(rvest_doc,"table")
str(test2)
```
And moving on.  
```{r}
test.table <-html_table(test)
head(test.table)
```
```{r}
str(test.table)
```
Pretty easy.  Note that html_nodes returns a list - I wonder if that would work for multiple tables on a page.  We shall see.  

But anyway, for a static page, we can use the selector gadget in Chrome to find an appropriate css term and then go from there.  Of course, this will only work with static pages - we really need to work on biomaRt tools.
