---
title: "Exploring Shiny"
author: "Bruce Cochrane"
date: "May 5, 2015"
output: html_document
---
Once again, we need to try to master this.  We're going to work through the [RStudio Tutorial](http://shiny/rstudio.com/tutorial).

## Lesson 1

I think we have it installed - we shall see 

```{r}
library(shiny)
```

###  The two components

So the idea is to make these two and put them together in their own directory as .R files.  One is called ui.R; the other is server.R  We'll start by doing so with the sample code in the tutorial (reproduced below for future reference)

#### The user interface

library(shiny)

 # Define UI for application that draws a histogram
shinyUI(fluidPage(

  # Application title
  titlePanel("Hello Shiny!"),

  # Sidebar with a slider input for the number of bins
  sidebarLayout(
    sidebarPanel(
      sliderInput("bins",
                  "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),

    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("distPlot")
    )
  )
))

#### The Server Instructions

library(shiny)

 # Define server logic required to draw a histogram
shinyServer(function(input, output) {

  # Expression that generates a histogram. The expression is
  # wrapped in a call to renderPlot to indicate that:
  #
  #  1) It is "reactive" and therefore should re-execute automatically
  #     when inputs change
  #  2) Its output type is a plot

  output$distPlot <- renderPlot({
    x    <- faithful[, 2]  # Old Faithful Geyser data
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    # draw the histogram with the specified number of bins
    hist(x, breaks = bins, col = 'darkgray', border = 'white')
  })


### Running the app
```{r}
runApp("./Shiny/App1")
```
OK so far.  And there's one other thing
```{r}
runApp("./Shiny/App1",display.mode='showcase')
```
Not sure what's up with that.

so there are a bunch of examples that run in showcase.  One follows:

```{r}
runExample('08_html')
```

## Lesson 2 - building a UI

So we will build the basics per the tutorial and put it in app2.

First, make the two wrappers

shinyUI(fluidPage(
  ))

Then create title and sidebar.  Note that default position for sidebar is left; can specify position=right
shinyServer(function(input,output){
  
})


OK.  So there are a bunch of tags we need to learn, but the basics of the ui are pretty logical.  We'll quit with this one and see if we can open a shiny-specific project that might be easier.


