---
title: <center>GWAS with genAbel</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---
Most of this taken and modified from [GenABEL Tutorial]()
if necessary
```{r}
setwd("./genAbel")
install.packages("GenABEL")
```
```{r}
library(GenABEL)
library(GenABEL.data)
```

```{r}
data(srdta)
```
The following gives the gender of each subject
```{r}
male(srdta)
```
Check the structure
```{r}
str(srdta)
```
And a map
```{r}
map(srdta)
```
I thinkt the following gives the phenotypic data
```{r}
head(phdata(srdta))
```
and the genotypic
```{r}
head(as.character(gtdata(srdta[,1:5])))
```
OK.  those are good.  We can subset in sort of the usual way:

```{r}
ssubs <-srdta[1:5,1:3]
summary(ssubs)
```
So for HW and call rate checking, we can work from the summary table with sumgt
```{r}
sumgt <-summary(gtdata(srdta))
```
That looks to be a nice simple data frame.  
```{r}
hist(sumgt$Pexact)
```
```{r}
hist(sumgt$CallRate)
```
### Doing an actual GWAS

Load another gwaa:
```{r}
data(ge03d2ex)
```
And although I don't like to, we can attach the phenotypic data
```{r}
attach(phdata(ge03d2ex))
```
#### Now a little bit of exploration of the phenotypes
```{r}
descriptives.trait(ge03d2ex)
```
separating by cases and controls.
```{r}
descriptives.trait(ge03d2ex,by=dm2)
```
So dm2 is presence or absence of diabetes.  

#### And the genotypes
```{r}
descriptives.marker(ge03d2ex)
```
There's a bunch of stuff about testing for HW, which we'll skip for now.

#### Running the GWAS
This is really clear.  We decide the classification variable (diabetes in this case) and then apply it to the data
```{r}
an0 <-qtscore(dm2,ge03d2ex,trait="binomial")
an0
```
Again, more stuff on testing, but we can do a 

#### Manhatten plot
```{r}
plot(an0)
```
```{r}
add.plot(an0,df="Pc1df",col=c("lightblue","lightgreen"))
```
Not a huge thing to worry about, but it's there.  More importantly

#### Finding the hits
```{r}
descriptives.scan(an0)
```
There's a lot more to explore, but for the purposes of teaching, the above may be enough.  What we'd like to have is more data!  A possible dryad source on [Bovine tuberculosis(http://datadryad.org/resource/doi:10.5061/dryad.519bm)]; really like the data for [this one on human eye color and HERC2](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2427174/).  However, it looks like human data aren't going to be there - [NIH site](http://www.ncbi.nlm.nih.gov/gap) requires IRB approval etc.
