---
title: <center>Chapter 7 - Incorporation of Stratification</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---

### Background

So in the chicken example, we have rather nice evidence for there being stratification - (although we haven't looked a t relatedness among individuals within populations.).  How we address that problem is dealt with in chapter 7 of the tutorial; for starters's we'll walk through some of the exercise.  However, to do so requires we generate some data following steps in chapter 5.

### Getting data1

```{r}
library(GenABEL)
library(GenABEL.data)
data(ge03d2ex)  
```
Now do the basic qc on this:
```{r}
qc1 <- check.marker(ge03d2ex, p.level=0)
```
```{r}
 data1 <- ge03d2ex[qc1$idok, qc1$snpok]
data1 <-Xfix(data1)
```
Attach the data
```{r}
attach(phdata(data1))
```
Now skip over the checks (we'll assume they're ok) and get the kinship matrix
```{r}
data1.gkin <-ibs(data1[,autosomal(data1)],weight="freq")
```
Do a pca on this just for grins
```{r}
data1.dist <-as.dist(0.5-data1.gkin)
data1.mds <-cmdscale(data1.dist)
plot(data1.mds)
```
So now what we'll do in this case is to use kmeans clustering, id the two clusters, and generate vectors of names of the two clusteds
```{r}
mk<-kmeans(data1.mds,centers=2,nstart=1000)
cl1 <-names(which(mk$cluster==1))
cl2 <-names(which(mk$cluster==2))
cl1
```
Note that in the above, the two could come out reversed; there is code to adjust for that in the tutorial.  we got it right (cl1 is outliers) so we will continue.

Now make a data set without the outliers
```{r}
data2 <-data1[cl2,]
```
Do QC
```{r}
qc2 <- check.marker(data2, hweids=(phdata(data2)$dm2==0), fdr=0.2)
```
```{r}
summary(qc2)
```
Now drop the failed markers (looks like all do to MAF)
```{r}
data2 <-data2[qc2$idok,qc2$snpok]
```
Now again there are the usual descriptives to run, but we will cut to the chase.

```{r}
detach(phdata(data1))
attach(phdata(data2))
```

```{r}
data2.qt <-qtscore(dm2,data2,trait="binomial")
```
And there's actually a lot more good stuff (e. g. on stratifying by other traits like bmi), but we'll save that for a future date.  Now we have all the variables we need to save (and to go to our chapter 7 workup)
```{r}
save(data1,cl1,data1.gkin,data2.qt,file="data1.RData")
```
### Doing our analysis

```{r}
## load(data1.RData) #necessary only if starting fresh from here.
```
Now we follow along, per pg. 167.

#### Do overall run and check lambda

```{r}
data1.qt <-qtscore(dm2,data1,trait="binomial")
lambda(data1.qt)
```
```{r}
cl1
```
That's ok.  Now we're going to isolate those cases as a population
```{r}
pop <-as.numeric(idnames(data1) %in% cl1)
table(pop,phdata(data1)$dm2)
```
And with that, we can do the stratification:
```{r}
data1.sa <-qtscore(dm2,data=data1,strata=pop, trait="binomial")
lambda(data1.sa)
```
We can plot the differences as follows
```{r}
par(mfrow=c(2,1))
plot(data1.qt)
plot(data1.sa)
```
Now this goes on and on, but I'm wondering if we could stop at this point and go back to the chickens.
```{r}
detach(phdata(data2))
```

