---
title: <center>Combining dataframes for genind</center>
author: <center>Bruce Cochrane</center>
date: <center>`r format(Sys.Date(),format="%B %d %Y")`</center>
output: 
    html_document:
       css: Chapter.css
       toc: yes
       toc_depth: 5
---
So with our function of the dat (do.call), we're going to try to make a single multipopulation data frame and do a complete genind analysis.  We'll start with our data from earlier.
```{r}
load("./Data/sampsnps.txt")
```
We'll also source our function to munge this
```{r}
source("./R/bmhap2df.R")
```

Now we can apply that function to the data
```{r}
genos <-lapply(snpdat.samp,bmhap2df)
```
Remind ourselves of the structure
```{r}
names(genos)
str(genos[[1]])
genos[[1]][1:5,1:5]
```
So there are two things to do.  First, we need to see if everything can be do.call(rbinded)
```{r}
genos.comb <-do.call(rbind.data.frame,genos)
```
So the problem with that is that the dimensions differ
```{r}
sapply(genos,dim)
```
So we need to find the intersection of all 11 sets of data
```{r}
snps <-lapply(genos,colnames)
```
Now, per R bloggers
```{r}
snp.common <-Reduce(intersect,(snps))
```
OK!  There are 267.  So do a quick test
```{r}
test <-genos[[1]]
test.2 <-test[,snp.common]
```
That's real easy.  So, we will do the following
```{r}
genos.2 <-lapply(genos, function (x) x[,snp.common])
sapply(genos.2,dim)
```
OK.  So now we should be able to do.call them
```{r}
genos.all <- do.call(rbind.data.frame, genos.2)
```
That looks excellent!  Now to make a population vector.

Here are the inputs:

```{r}
pops <-names(genos.2)
popsize <-sapply(genos.2,dim)[1,]
pops;popsize
```
```{r}
popvec <-unlist(lapply(1:11, function (x) rep(pops[x],popsize[x]))
```
That looks good.  So to make the genind, we want to do the following
```{r}
genos.gen <-df2genind(genos.all,pop=popvec)
```
```{r}
out <-summary(genos.gen)
```
And this looks gorgeous!!  So just to complete the basic stats stuff from the nescent procedure, we'll try the following:

```{r}
library(hierfstat)
library(HardyWeinberg)
```
```{r}
genos.hf <-genind2hierfstat(genos.gen)
basicstat <- basic.stats(genos.hf, diploid = TRUE, digits = 2) 
```
That looks like fun.
```{r}
str(basicstat)
```
And do a pca:
```{r}
x <-indpca(genos.hf)
plot(x,cex=0.5,col=c(1:11))
```
This is an absolutely fascinating outcome.  We need to work on the visualization, and also to test it with other regions.  But we now have the workflow:

1.  Do the biomaRt query based on the combined database
2.  Split into a list
3.  apply bmhap2df to the list
4.  Find the common snps as above and subset
5.  do.call to make a dataframe
6.  Make the population vector
7.  Make a genind.
8.  Make a hierfstat object or whatever.

Two last tests
```{r}
library(multiNe)
```
```{r}
HENe(genos.gen)
```
That function is really sketchy - even with all of this work, we still only have an Ne of 600 or so.

Test for Hardy Weinberg

```{r}
genos.loc <-genind2loci(genos.gen)
hw.test(genos.loc,B=100)
```
Lots of non hw here.  But remember - this is combined data.

